---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Author: I. Diaz

## Loading and preprocessing the data

Information is zipped into a file, so first is necesary to unzip the data and then save de dataset.
```{r, echo=TRUE}
#unzip de file
unzip("activity.zip")

#saving dataset as "dataset"
dataset <- read.csv("activity.csv")
head(dataset)
summary(dataset)
```

Setting date column to date format
```{r, echo=TRUE}
dataset$date <- as.Date(as.character(dataset$date), format = "%Y-%m-%d")
head(dataset)
```

Exploring data without NAs
```{r, echo=TRUE}
datasetna <- dataset[complete.cases(dataset),]
head(datasetna)
```

## What is mean total number of steps taken per day?
First of all, I'll aggregate datasetna by summing steps by date and save the results to "aggdata"  in order to prepare de data to build a histogra and calculate de mean and median.
```{r, echo=TRUE}
aggdata <- aggregate(steps ~ date, data = datasetna, FUN = sum)
head(aggdata, 10)
```

Ploting a histogram

```{r, echo=TRUE}
library(ggplot2)
hist1 <- ggplot(aggdata, aes(x = steps)) + geom_histogram(fill = "darkred", binwidth=2000) + labs(title = "Total Steps per day", x = "Steps", y = "Count")

hist1
```

Calculating the mean and median
```{r, echo=TRUE}
paste("the mean is ", mean(aggdata$steps))
paste("the median is ", median(aggdata$steps))
```

## What is the average daily activity pattern?
In order to get the data prepared and build the timeserie is necesary to get the average of steps by intervale. I've used the aggregate function to have this.
```{r, echo=TRUE}
avrgsteps <- aggregate(steps ~ interval, data = datasetna, FUN = mean)
head(avrgsteps, 10)
```

Build the time series using
```{r, echo=TRUE}
timeserie <- ggplot(avrgsteps, aes(x = interval, y = steps)) + geom_line(color = "darkred") + labs(title = "Timeserie: Steps by Interval", x = "5-min Interval", y = "Steps Avrg")

timeserie
```

To calculate the max value I'll use de previous set "avrgsteps" I created, because it contains the average number of step for every 5-min interval.

```{r, echo=TRUE}
avrgsteps[which(avrgsteps$steps == max(avrgsteps$steps)),]
```

## Imputing missing values
Accordingly to the summary function previously used to the original dataset, the number of NAs is 2304.

I'll try to process the empty values in order to check if this take effect to the final result. NAs will be replace with the mean.
```{r, echo=TRUE}
#calculating the mean value
meanvalue <- mean(datasetna$steps)
meanvalue

#replacing NAs with the mean in the "steps" column
datasetnona <- dataset
nas <- is.na(dataset$steps)
datasetnona$steps[nas] <- meanvalue
summary(datasetnona)
```

Build a histogram
```{r, echo=TRUE}
aggdata2 <- aggregate(steps ~ date, data = datasetnona, FUN = sum)
hist2 <- ggplot(aggdata2, aes(x = steps)) + geom_histogram(fill = "darkblue", binwidth=2000) + labs(title = "Total Steps per day", x = "Steps", y = "Count")

hist2
```

Calculate the mean and the median of the de data aggregated

```{r, echo=TRUE}
paste("the mean is ", mean(aggdata2$steps))
paste("the median is ", median(aggdata2$steps))
```

If we see there is no significant impact of the in the values repect the original dataset, just in case of the median because of the 2304 NAs that not exist anymore and center value changed.

## Are there differences in activity patterns between weekdays and weekends?

I'll create a new variable call "weekd" and call the weekday using the weekday() function. Then if weekd is not weekend store 1 in the "weekdclasif" variable and 0 if it's weekend. 
```{r, echo=TRUE}
datasetnona2 <- datasetnona
datasetnona2["weekd"] <- weekdays(datasetnona2$date)
# assign weekdays to 1 and weekend to 0 and save the values to "weekdclasif"
datasetnona2$weekdclasif <- ifelse(datasetnona2$weekd == "Saturday"  | datasetnona2$weekd == "Sunday", "weekend", "weekday")
tail(datasetnona2,5)
```

Is necessary to prepare the dataset and then build the plot in order to compare how is the variation of the steps on weekdays and weekend.
```{r, echo=TRUE}
aggdata3 <- aggregate(datasetnona2$steps, by=list(datasetnona2$weekdclasif, datasetnona2$interval), mean)
names(aggdata3)[1]  = "weekd"
names(aggdata3)[2] = "interval"
names(aggdata3)[3] = "meansteps"

head(aggdata3)
```
Buils de plot using ggplot

```{r, echo=TRUE}
timeserie2 <- ggplot(data = aggdata3, aes(x=interval, y=meansteps, color=weekd)) + facet_grid(weekd ~ .) + geom_line() + labs(title = "Steps by 5-min Interval, Classified by Weekday", x= "Interval", y="Avrg Steps")

timeserie2
```

Looks like during the weekend is more common to have more steps.